Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-02-06 WSSSPE3/report/wg_appendix_publishing_SW.tex
changes to changes in later sections

Create the first version of the web page to be displayed on the Software
Sustainability Institute's website. %: \url{http://www.software.ac.uk}. \choinote{already appeared above)
We expect the page to be live in early January of 2016.

@@

Steven R.\ Brandt has created a first version of the page, and it is in the process
of being posted on the Software
Sustainability Institute's website. %: \url{http://www.software.ac.uk}.  \choinote{already appeared above)
Neil Chue Hong will take responsibility for the page once it is up.

>>>>

Create the first version of the web page to be displayed on the Software
Sustainability Institute's website: \url{http://www.software.ac.uk}.
% \choinote{already appeared above) /katznote{but important to have it here for poeple who are scanning/looking at this part of each section}
We expect the page to be live in early January of 2016.

@@

Steven R.\ Brandt has created a first version of the page, and it is in the process
of being posted on the Software
Sustainability Institute's website: \url{http://www.software.ac.uk}. % \choinote{already appeared above)
Neil Chue Hong will take responsibility for the page once it is up.

### 2016-02-06 WSSSPE3/report/wg_appendix_software_credit.tex
Appendix J: Remove an unintended change in comment.
changes to changes in later sections

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Group Members}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% g

@@

Initial discussions mainly focused on both various mechanisms for, and the
philosophical approach behind, crediting software in scientific papers. These
began with proposals for various ways to credit software (or other research
products including data) that contributed more significantly than a generic
citation, including:
\begin{itemize}

@@

\item Mozilla Science Lab has recently introduced Contributorship Badges for
Science~\cite{Mozilla_badges}, which provides a badge---associated with
ORCID~\cite{orcid}---for identifying author contributions in a paper based on the taxonomy
outlined in Project CRediT.

@@

During the first day of WSSSPE3 breakout sessions, the group discussed and deliberated over the
Entertainment Identifier Registry (EIDR)~\cite{EIDR} as a potential model for
scientific software. That system assigns unique Digital Object Identifiers
(DOIs)---the same system used for scientific publications---to all content
(e.g., movies, television shows) and contributors, along with relevant metadata.
One important use of the EIDR system is to track rights and credits for contributors
to entertainment works in order to distribute revenues---similar to the proposed
transitive credit concept.

@@

On the second day of WSSSPE3, the group decided to put aside the
taxonomy of contributions and focus on software citations to ensure developers
receive credit (regardless of contribution). Eventually, once 
software citations are standardized, the goal would be to return to establishing different
roles or contributions for this credit. Following this decision, the group
identified two necessary actions to move forward:
\begin{enumerate}

@@

Although far from a standard practice, examples of citing software in
publications can be found in various scientific communities---notably,
representative samples can be found in astronomy~\cite{astronomy_SW_examples}
and biology~\cite{Howison2015}. The group recommended collecting similar
examples from other communities, and then developing a set of software citation principles
document in concert with the FORCE11 Software Citation Working Group (see
\S\ref{SC:plan} for more details), following the model of the FORCE11 Data
Citation Principles document~\cite{DataCitation2014}.

The group further discussed briefly whether software used directly in a
publication---whether to perform simulation or analysis, or as a dependency for
new software to be developed---should be distinguished from other references due to
the dependence of the study on these research artifacts. Suggestions included a
separate list of citations (with DOIs) for software and other research objects
that serve this sort of ``vital'' role. Similar recommendations were made by the
credit breakout group at WSSSPE2~\cite{WSSSPE2}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Group Members}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Initial discussions focused on both various mechanisms for, and the
philosophical approach behind, crediting software in scientific papers. These
began with proposals for various ways to credit software (or other research
products including data) that contributed more significantly than a generic
citation, including:
\begin{itemize}

@@

\item Mozilla Science Lab' recently introduced Contributorship Badges for
Science~\cite{Mozilla_badges}, which provide a badge---associated with an
ORCID~\cite{orcid}---that recognizes author contributions using the taxonomy
outlined in Project CRediT.

@@

In the breakout sessions on the first day of WSSSPE3, the group discussed and deliberated over the
Entertainment Identifier Registry (EIDR)~\cite{EIDR} as a potential model for
scientific software. That system assigns unique Digital Object Identifiers
(DOIs)---the same system used for scientific publications---to all content
(e.g., movies, television shows) and contributors, along with relevant metadata.
One important use of the EIDR system is to track rights and credits for contributors
to entertainment works in order to distribute revenues---similar to the proposed
transitive credit concept.

@@

On the second day of WSSSPE3, the group decided to put aside the
taxonomy of contributions and focus on software citations to ensure developers
receive credit (regardless of contribution). Eventually, once 
software citations are standardized, the goal would be to return to establishing different
roles\slash contributions for this credit. Following this decision, the group
identified two necessary actions to move forward:
\begin{enumerate}

@@

Although far from a standard practice, examples of citing software in
publications can be found in various scientific communities---notably,
representative samples can be found in astronomy~\cite{astronomy_SW_examples}
and biology~\cite{Howison2015}. The group recommended collecting similar
examples from other communities, and then developing a software citation principles
document in concert with the FORCE11 Software Citation Working Group (see
\S\ref{SC:plan} for more details), following the model of the FORCE11 Data
Citation Principles document~\cite{DataCitation2014}.

The group further discussed briefly whether software used directly in a
publication---whether to perform simulation or analysis, or as a dependency for
newly developed software---should be distinguished from other references due to
the dependence of the study on these research artifacts. Suggestions included a
separate list of citations (with DOIs) for software and other research objects
that serve this sort of ``vital'' role. Similar recommendations were made by the
credit breakout group at WSSSPE2~\cite{WSSSPE2}.

### 2016-02-06 WSSSPE3/report/wg_main_publishing_SW.tex
changes to changes in later sections

\item\textbf{Software Credit (\S\ref{sec:software-credit})}: Since these forms of
publishing must make their sources explicit in order to execute, they should be
easier to trace even if appropriately worded credit for software is not
provided. In addition, some  notebooks, i.e., active or executable papers,
(see Appendix~\ref{sec:appendix_publishing_SW} for some concrete examples) 
make it possible to provide or define
additional metadata to make the tracing of credit clearer. Finally, attributions
could be added to citations to identify whether a paper extends a result,
verifies it, contradicts it, etc.

>>>>

\item\textbf{Software Credit (\S\ref{sec:software-credit})}: Since these forms of
publishing must make their sources explicit in order to execute, they should be
easier to trace even if appropriately worded credit for software is not
provided. In addition, they
make it possible to provide or define
additional metadata to make the tracing of credit clearer. Finally, attributions
could be added to citations to identify whether a paper extends a result,
verifies it, contradicts it, etc.
